---
title: "DAG"
author: "Fiona Grimm"
date: "29/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggdag)
theme_set(theme_dag())
```

# Directed acyclic graph (DAG)
DAG showing known causes of confounding when comparing
hospital mortality risk across study populations from 
before and during the pandemic

```{r}
hip_dag <- dagify(COMP ~ PROC,
                  PROC ~ FRACTYPE, 
                  QUAL ~ PAND,
                  AGE ~ PAND,
                  SEX ~ PAND,
                  COVID ~ PAND,
                  PROC ~ HEALTH + FRACTYPE + PAND,
                  DEATH ~ HEALTH + QUAL + COMP + AGE + COVID + SEX,

#create labels for variables ("\n" denotes a space)
       labels = c("DEATH" = "Death", 
                  "HEALTH" = "Existing\nconditions",
                  "AGE" = "Age",
                  "SEX" = "Sex",
                  "COVID" = "COVID-19",
                  "COMP" = "Complications",
                  "PROC" = "Procedure type",
                  "FRACTYPE" = "Fracture type",
                  "PAND" = "Pandemic",
                  "QUAL" = "Quality of care"),

#identify exposure and outcome     
                  exposure = "QUAL",
                  outcome = "DEATH",
                  latent = c("PAND", "FRACTYPE", "COMP"))
```

Notes:

* Confounder: common cause of both the exposure and the outcome
* Mediator: caused by the exposure and in turn causes the outcome (i.e. falls on a causal path between the exposure and outcome)
* Competing exposure: cause of the outcome that is neither caused by nor causes the exposure

A confounding path (or indirect path or backdoor path) between an exposure and an outcome is a path that passes against the flow of causality through one or more confounders.

A collider path between an exposure and an outcome is a path that passes against the flow of causality through one or more colliders, i.e. nodes that receive two ormore arcs.

```{r}
#plot DAG using the ggdag function
ggdag(hip_dag, text = FALSE, use_labels = "label", edge_type="link")
```


```{r}
ggsave("hipfrac_mortality_dag.png", width = 9, heigh = 5)

```


# Open paths 

```{r}
ggdag_paths(hip_dag, text = FALSE, use_labels = "label", shadow = TRUE, 

#directed=false indicates all paths, regardless of direction            
            directed=FALSE,  

#changing node/text sizes for aesthetics             
            node_size = 8,   text_size = 2, stylized = TRUE)
```


# Directed paths

```{r}
ggdag_paths(hip_dag, text = FALSE, use_labels = "label", shadow = TRUE, 

#directed=true indicates directed paths            
            directed=TRUE,  
            node_size = 10,   text_size = 2.8, stylized = TRUE)
```

# Minimally sufficient set

```{r}


ggdag_adjustment_set(hip_dag, text = FALSE, use_labels = "label", 
                     shadow = TRUE, node_size = 10,   text_size = 2.8, stylized = TRUE)
```


```{r}
ggsave("hipfrac_mortality_minadjustset.png", width = 9, heigh = 5)

```
